---
title: "Computer Assignment 4b Costello"
author: "Rebecca Costello"
date: "9/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Ties, Rebecca and Niels

```{r packages}
library(foreign)
library(tidyverse)
library(ggdag)
library(dplyr)
library(tinytex)
library(jtools)
library(huxtable)
library(summarytools)
library(ggstance)
library(pwr)
library(knitr)
library(lemon)
library(AER)
library(rdrobust)
library(lubridate)

```

```{r layout}
knit_print.data.frame <- lemon_print
st_options(plain.ascii = FALSE, style = "rmarkdown")
st_css()
```

```{r data}
theUrl_ca4b <- "https://surfdrive.surf.nl/files/index.php/s/JAo4jnvf042xQXt/download"
fertility <- read.dta (file = theUrl_ca4b)
```

1. Mechanism
Y(0,i)= A baby will not be concieved
Y(1,i)= A baby will be conceived and born whereby the mother recieves the cash benefit of 2,500eu

Why fertility may be affected by the child benefit:
-Having a child is an expensive choice and the benefit may ease the cost burden of having a child, allowing the parents to provide a slightly better life.
-People may be incentivised by the child benefit to pay for things they need.

2. Descriptive Statistics
(a) What is the average number of abortions per month in your data sample for Spain (representing 88
percent of national abortions, see page 165 of the paper)?

```{r avg abortions}
abortpermth <- fertility %>% group_by(month(fertility$abortions), mday(fertility$abortions)) %>% tally()
mean(abortpermth)
```













