---
title: "Computer Assignment 4b Costello"
author: "Rebecca Costello"
date: "9/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Ties, Rebecca and Niels

```{r packages}
options(repos="https://cran.rstudio.com")
install.packages("jtools")
install.packages("huxtable")
install.packages("ggstance")
install.packages("summarytools")
install.packages("pwr")
install.packages("knitr")
install.packages("lemon")
install.packages("tinytex")
install.packages("AER")
install.packages("rdrobust")
install.packages("lubridate")

```

```{r packages II}
install.packages(("tidyverse"))
```

```{r packages III}
install.packages("rdrobust")
install.packages("ggdag")
```


```{r packages}
library(foreign)
library(tidyverse)
library(ggdag)
library(dplyr)
library(tinytex)
library(jtools)
library(huxtable)
library(summarytools)
library(ggstance)
library(pwr)
library(knitr)
library(lemon)
library(AER)
library(rdrobust)
library(lubridate)

```

```{r layout}
knit_print.data.frame <- lemon_print
st_options(plain.ascii = FALSE, style = "rmarkdown")
st_css()
```

```{r data}
theUrl_ca4b <- "https://surfdrive.surf.nl/files/index.php/s/JAo4jnvf042xQXt/download"
fertility <- read.dta (file = theUrl_ca4b)
```

1. Mechanism
Y(0,i)= A baby will not be concieved
Y(1,i)= A baby will be conceived and born whereby the mother recieves the cash benefit of 2,500eu

Why fertility may be affected by the child benefit:
-Having a child is an expensive choice and the benefit may ease the cost burden of having a child, providing more disposable income, allowing the parents to provide a slightly better life.
-People may be incentivised by the child benefit to pay for things they need.

2. Descriptive Statistics
(a) What is the average number of abortions per month in your data sample for Spain (representing 88
percent of national abortions, see page 165 of the paper)?

```{r avg abortions}
summary(fertility$abortions)

```




```{r}
bycalmonth <- fertility %>% group_by(calendar_month)
mean(bycalmonth$abortions)
```


(b) In percentage terms, how much higher would the number of births be if there were no abortions (number
of births is directly based on number of conceptions, number of births is based on all of Spain, and
assume that the national estimate for the number of abortions is 1.14 times the average reported for
the sample of 14 regions).

```{r}
bycalmonth <- fertility %>% group_by(calendar_month) 
mean(bycalmonth$conceptions)

```
```{r conceptions percent }
summary(fertility$conceptions)

```

(c) What is the complete calendar time window for which you have data? To create a date variable, you
can do the following:

```{r cal time window}
fertility$date1 <- fertility$calendar_year+(fertility$calendar_month/12)
fertility$date1 <- fertility$date1-(1/12)
fertility$date1 <- as.yearmon(fertility$date1)

fertility$date1 <- as.Date(fertility$date1)
min(fertility$date1,na.rm=TRUE)
max(fertility$date1,na.rm=TRUE)


```
The time window is from the 1st of April 2001 until 1st of December 2009.

(d) Using a scatterplot and the date-variable date, show the general trend in the number of abortions over
time. Make sure the vertical axis ranges from 0 to 10,000.

```{r}
plot(fertility$date1, fertility$abortions,
     xlab="Year", ylab="Avg number of abortions",
     ylim = c(0, 10000))
abline(lm(fertility$abortions ~ fertility$date1))
```

3. Graphical evidence

(a) Show how the number of conceptions changes around the cutoff date, taking month of conception as
the running variable.

running variable should be 

```{r}
monconcprange <- fertility$month_of_conception
monconcprange <- range(-40, 29)
```


```{r}
is.na(monconcprange)
```
```{r}
with(fertility, rdplot(y = conceptions, x = month_of_conception, p=1, c=0, nbins = 35,
                        title="Discontinuity conception",
                        x.label="month of conception relative to July 1, 2007",
                        y.label="number of conceptions"))
``` 

(b) What is the treatment effect in terms of conceptions? Infer this from eyeballing the discontinuity at
the cutoff point (so a rough idea of the difference is sufficient here; we will derive exact answers later
on when you run regressions). Derive the approximate percentage change.

There is a spike in conceptions around the cutoff point (with a increase leading up to it), and then a decline as conceptions return to the average. This shows self_selection into treatment. It goes from 40000 to 44000, indicating a 10 per cent increase.


(c) Do the same thing as asked under (a) and (b), but now for the number of abortions and with month of
birth as running variable. The figure you are going to plot is a nice illustration of the importance of the
functional form that you impose on the relationship between the running variable and the outcome.

```{r}
with(fertility, rdplot(y = abortions, x = month_of_birth, p=1, c=0, nbins = 35,
                        title="Discontinuity abortions",
                        x.label="month of birth relative to July 1, 2007",
                        y.label="number of abortions"))
``` 


Abortions can be seen to slowly decrease after the cutoff point of the policy being introduced. 


4. Test of treatment effect
1. Treatment effect on conception (ln_conceptions)
```{r naive reg conception}
reg4b1 <- lm(ln_conceptions ~ treatment_dummy, data=fertility)
summ(reg4b1)
```


2. Treatment effect on abortions (ln_abortions)

```{r naive reg abortions}
reg4b2 <- lm(ln_abortions ~ treatment_dummy, data=fertility)
summ(reg4b2)
```












